---
title: "R Markdown Example"
author: "Simon Fontaine"
date: "20/08/2021"
output: 
  html_document:
    theme: cosmo
    highlight: kate
    toc: True
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#")
```

## R Markdown

You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure}
plot(pressure)
```

## Displaying data frames

Markdown does some pretty formatting of data frames into tables:

```{r df}
knitr::kable(mtcars)
```

## Chunk options

A R chunk looks like this (with triple backward ticks):
```
``{r label, options1=..., option2=...}
# R Code goes here
``
```

You can look [here](https://yihui.org/knitr/options/)
for a complete list, but here some of the most relevant:

- `echo=T/F` to display the code or not (the output will still appear)
- `eval=T/F` to evaluate the code or note
- `warning=T/F` and `error=T/F` to suppress warning and error messages
- `comment="##"` what will appear at the beginning of output lines
- `cache=T/F` whether to cache results for future compiliation (if they are modified, they will be re-run)
- `fig.x`: a bunch of options for figures

You can also set options for the whole document using

```{r eval=F}
knitr::opts_chunk$set(echo = TRUE, comment = "#>#")
```

## Markdown syntax

See [this markdown cheat sheet](https://www.markdown-cheatsheet.com/) 
for a list!


### Header

Header for this file:
```{yaml}
---
title: "Example"
author: "Simon Fontaine"
date: "20/08/2021"
output: 
  html_document:        # options for html output 
    theme: cosmo        # theme
    highlight: kate     # syntax highlighting
    toc: True           # add a table of content
---
```
See [themes](https://www.datadreaming.org/post/r-markdown-theme-gallery/) and [syntax highlighting styles](https://eranraviv.com/syntax-highlighting-style-in-rmarkdown/).

### LaTeX Math 


You can write math inline: `$\mathbb P[X=x]=\tfrac 12$` gives $\mathbb P[X=x]=\tfrac 12$

You can write math in display (`$$ ... $$` also works):


```{latex}
\begin{equation}
  \sum_{i=1}^n (2\pi)^{i} \int f_i(x) \,\text{d} x
\end{equation}
```
\begin{equation}
  \sum_{i=1}^n (2\pi)^{i} \int f_i(x) \,\text{d} x
\end{equation}

To get aligned equations, you need to use `align` environments (`aligned` environments within `$$ ... $$` also works)

```{latex}
\begin{align}
  f_X(x) 
  &= \int f_{X,Y}(x,y) \text{d} y \\
  &= \int f_Y(y)f_{X\mid Y}(x\mid y) \,\text{d} y 
\end{align}
```

\begin{align}
  f_X(x) 
  &= \int f_{X,Y}(x,y) \text{d} y \\
  &= \int f_Y(y)f_{X\mid Y}(x\mid y) \,\text{d} y 
\end{align}




### Pros and cons

Pros:

- Much easier/readable syntax than LaTeX
- Seamless integration of code, output, and math

Cons:

- Math can be a little trickier (cannot define macros in html, but there's a way for pdf output!)
- More limited than LaTeX

