---
title: "Reproducible and Distributable Code"
author: "Simon Fontaine"
date: "19/08/2021"
output: 
  html_document:
    theme: cosmo
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

A collection of dos and don'ts, tips and tricks in programming to make your code
more reproducible, readable and reusable distributable!

We'll be working in R, but most of these apply to any language!

## Dos and Don'ts

<!-- ======================================================================= -->
### DON'T: hard-code parameters, dimensions, constants

Example: generating a $100\times 10$ matrix of uniform covariates
```{r}
X = matrix(runif(100*10), nrow=100, ncol=10)
```

### DO: extract in variables
```{r}
n = 100
p = 10
X = matrix(runif(n*p), nrow=n, ncol=p)
```

<!-- ======================================================================= -->
### DON'T: copy/paste
Example: generate data from different parameters, then fit a simple linear regression.
```{r}
set.seed(0)
x = runif(n)
y = 1 + 2*x + rnorm(n)
fit = lm(y ~ x)
print(coef(fit))
x = runif(n)
y = -1 + 1*x + rnorm(n)
fit = lm(y ~ x)
print(coef(fit))
```

### DO: extract functions

With the same number of lines, we now have reusable code!

```{r}
generate_data_and_fit = function(n, a, b){
  x = runif(n)
  y = a + b*x + rnorm(n)
  fit = lm(y~x)
  return(fit)
}
set.seed(0)
fit1 = generate_data_and_fit(n, 1, 2)
print(coef(fit1))
fit2 = generate_data_and_fit(n, -1, 1)
print(coef(fit2))
```


<!-- ======================================================================= -->
### DON'T: use global variables within functions

While your logic might still produce the correct output, this often lead to errors
and hard to understand code.

```{r}
set.seed(0)
x = runif(n)
y = 1 + 2*x + rnorm(n)
fit_model = function(){
  return(coef(lm(y ~ x)))
}
fit_model()
y = 1 + 2*x + rnorm(n)
fit_model()
```


Similarly, watch out for local/global variables with the same name; this 
produces confusing code to read and may not behave as you expect.

```{r}
x = 3
fn = function(){
  x = 4
}
print(x)
fn()
print(x)
```

### DO: pass everything as arguments

We get the same output as above, but this is much clearer now. This is especially
important if there is more space between calls.

```{r}
set.seed(0)
x = runif(n)
y = 1 + 2*x + rnorm(n)
fit_model = function(x, y){
  return(coef(lm(y ~ x)))
}
fit_model(x, y)
y = 1 + 2*x + rnorm(n)
fit_model(x, y)
```



<!-- ======================================================================= -->
### DON'T: use comment/uncomment for multiple calls

```{r}
set.seed(0)
a = 1; b = 2
# a = -1; b = 1
x = runif(n)
y = a + b*x + rnorm(n)
fit = lm(y~x)
print(coef(fit))
```

(comment/uncomment in the editor to switch cases)

```{r}
# a = 1; b = 2
a = -1; b = 1
x = runif(n)
y = a + b*x + rnorm(n)
fit = lm(y~x)
print(coef(fit))
```


For a single line and only two cases, this doesn't seem too bad, but:

- Imagine you have a few lines to switch at multiple locations; it becomes hard to remember what to switch
- Imagine someone else has to run your code

### DO: extract functions

```{r}
generate_data_and_fit = function(n, a, b){
  x = runif(n)
  y = a + b*x + rnorm(n)
  fit = lm(y~x)
  return(fit)
}
set.seed(0)
coef(generate_data_and_fit(n, 1, 2))
coef(generate_data_and_fit(n, -1, 1))
```



<!-- ======================================================================= -->
### DON'T: write big functions

Example: compare estimates from two models

```{r}
generate_data_and_compare = function(n, a, b){
  x = runif(n)
  y = a + b*x + rnorm(n)
  slr = lm(y ~ x)
  intercept_slr = coef(slr)[1]
  mean = lm(y ~ 1)
  intercept_mean = coef(mean)[1]
  return(intercept_slr - intercept_mean)
}
set.seed(0)
generate_data_and_compare(n, 1, 2)
```


### DO: split into steps

```{r}
generate_data = function(n, a, b){
  x = runif(n)
  y = a + b*x + rnorm(n)
  return(data.frame(x, y))
}
get_intercept = function(model){
  return(coef(model)[1])
}
compare_models = function(df){
  slr = lm(y~x, data=df)
  mean = lm(y~1, data=df)
  return(get_intercept(slr) - get_intercept(mean))
}
set.seed(0)
df = generate_data(n, 1, 2)
compare_models(df)
```




<!-- ======================================================================= -->
### DON'T: chain too many operations

```{r}
set.seed(0)
x = runif(n)
y = 1 + 2*x + rnorm(n)
coef(lm(y~x))[1] - coef(lm(y~1))[1]
```


### DO: use intermediary assignments

```{r}
slr = lm(y ~ x)
intercept_slr = coef(slr)[1]
mean = lm(y ~ 1)
intercept_mean = coef(mean)[1]
intercept_slr - intercept_mean
```




<!-- ======================================================================= -->
### DO: spell out named arguments

R (& Python) allows for both ordered unnamed arguments and named arguments.
It's hard to spell out "rule", but

- optional arguments (those with defaults) should be named
- large numbers of arguments should be named (possibly beyond the first few ones that are obvious)
- this mostly applies to larger functions, e.g., those in packages

```{r}
lm(y~x, data=df, qr=FALSE)
```



<!-- ======================================================================= -->
## Tips and tricks
<!-- ======================================================================= -->
### Variable and function name

- use meaningful names, but keep it short
- be consistent
<!-- ======================================================================= -->
### Use visual aides

Increase readability by:

- using consistent indentation
- using empty lines and/or comments to separate steps
- using consistent delimiters and spacing

This is unreadable:

```{r}
generate_data = function(n, a, b){
x = runif(n)
y = a + b*x + rnorm(n)
return(data.frame(x, y))}
get_intercept = function(model){return(coef(model)[1])}
compare_models = function(df){
slr = lm(y~x, data=df)
mean = lm(y~1, data=df)
return(get_intercept(slr) - get_intercept(mean))}
set.seed(0)
df = generate_data(n, 1, 2)
compare_models(df)
```

This is much clearer

```{r}
# helper functions
generate_data = function(n, a, b){
  x = runif(n)
  y = a + b * x + rnorm(n)
  return(data.frame(x, y))
}
get_intercept = function(model){
  return(coef(model)[1])
}
compare_models = function(df){
  slr = lm(y ~ x, data=df)
  mean = lm(y ~ 1, data=df)
  return(get_intercept(slr) - get_intercept(mean))
}

# run comparison
set.seed(0)
df = generate_data(n, 1, 2)
compare_models(df)
```

<!-- ======================================================================= -->
### Comment what you are doing

- If you use meaningful names for variables and functions, you don't need too much comments
- At the same time, it's rare that there are too much comments
- Write comments for steps that are not obvious

This is unnecessary unless you are writing documentation:

```{r}
# function to generate data
# single covariate with uniform distribution on [0,1]
# outcome y=a+bx+e for some N(0,1) error
generate_data = function(n, a, b){
  x = runif(n)
  y = a + b * x + rnorm(n)
  return(data.frame(x, y))
}
# function to extract the intercept from a model
get_intercept = function(model){
  return(coef(model)[1])
}
# function that compute a OLS model using the covariate
# and a mean model and returns the difference in estimated intercept
compare_models = function(df){
  slr = lm(y ~ x, data=df)
  mean = lm(y ~ 1, data=df)
  return(get_intercept(slr) - get_intercept(mean))
}

# set seed for replicability
set.seed(0)
# generate the data
df = generate_data(n, 1, 2)
# get difference in intercept
compare_models(df)
```

<!-- ======================================================================= -->
### Be as independent of your setup as possible

If you need to read/write files:

- use relative paths
- have a few lines at the top that are easily changed

Don't do this:
```{r eval=FALSE}
# line 100
read.csv("/home/simon/Documents/prep_workshop_2021/data1.csv")

# line 200
read.csv("/home/simon/Documents/prep_workshop_2021/data2.csv")

# line 300
write.csv("/home/simon/Documents/prep_workshop_2021/results.csv")
```

Do this:
```{r eval=FALSE}
# line 1
setwd("/home/simon/Documents/prep_workshop_2021/")

# line 101
read.csv("data1.csv")

# line 201
read.csv("data2.csv")

# line 301
write.csv("results.csv")
```

Or this:
```{r eval=FALSE}
# line 1
PATH = "/home/simon/Documents/prep_workshop_2021/"

# line 101
read.csv(paste0(PATH, "data1.csv"))

# line 201
read.csv(paste0(PATH, "data2.csv"))

# line 301
write.csv(paste0(PATH, "results.csv"))
```



<!-- ======================================================================= -->
### Plan ahead

It can be useful to draft the outline of your code before actually writing it:

- using comments
- using placeholder functions

This can help you think about what steps and functions you will need; what
is a parameter, what should be the arguments, etc.

```{r eval=FALSE}
# setup section


# parameters


# helper functions
generate_data = function(n, a, b){
  return(NULL)
}
get_comparison = function(df){
  return(NULL)
}


# run experiment


# store results

```



<!-- ======================================================================= -->
### Misc

- Split into different files as much as possible
- Refresh your environment once in a while and re-run the whole file to check if
you wrote things dependent on global variables that are no longer defined
- Use seeding whenever you randomize things




